---
title: "Final_Project"
author: "Mengqi Zhu"
date: "2018/11/29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr) 
library(lme4)
library(janitor)
library(corrplot)
library(ggplot2)
library(broom)
library(mice)

```

```{r}
smoke <- read.csv(file = 'frmgham2.csv') %>% 
  clean_names()
```



```{r}
#checking missing values (5% rule)
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(smoke,2,pMiss) #2 indicates columns
#Remove hdlc, ldlc, glucose, and bpmeds for having more than 5% of missing values.
```
It is okay as hdlc, ldlc are highly correlated with totchol, glucose is highy correlated with diabetes, bpmeds is highly correlated with sysbp and diabp. Therefore it won't lose much information to just drop these features.

#Part1
# Question 1
Figure 1 shows that as individuals age, the likelihood that they are smoking decreases. We can see that when we breaking individuals down by sex, it appears that the overall trend is the same between sexes with males having an overall higher likelihood of being smokers as age increases. 

```{r}

smoke %>%
  ggplot(aes(age, cursmoke)) +
  geom_jitter(height = 0.1, size = 0.5) +
  geom_smooth(method = "loess") +
  ggtitle("Figure 1: Current Smoking Status across Age") 

#BY SEX
smoke %>%
  mutate(sex = as.factor(sex)) %>%
  ggplot(aes(age, cursmoke, group = sex, color = sex)) +
  geom_jitter(height = 0.1, size = 0.5) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Figure 2: Current Smoking Status across Age") 
```

```{r}
smoke_vs1 = smoke %>% 
  filter(period==1) %>% 
  dplyr::select(c(randid,sex,age,cursmoke,totchol,bmi,heartrte,educ,diabp,sysbp,diabetes,prevap,prevchd,prevmi,prevstrk,prevhyp)) %>% 
  mutate(cursmoke = as.factor(cursmoke), sex=as.factor(sex), diabetes=as.factor(diabetes)) %>% 
  na.omit()
```

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+prevhyp, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
prevhyp is not confounder for sex and age with cursmoke
```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+prevstrk, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
prevstrk is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+prevmi, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
prevmi is not confounder for sex and age with cursmoke
```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+prevchd, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
prevchd is not confounder for sex and age with cursmoke
```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+prevap, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
prevap is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+totchol, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
totchol is not confounder for sex and age with cursmoke



```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+sysbp, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
sysbp is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+diabp, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
diabp is is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+bmi, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
bmi is is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+heartrte, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
heartrte is is not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+educ, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
educ is  not confounder for sex and age with cursmoke

```{r}
glm1 <- glm(cursmoke ~ age+sex, data=smoke_vs1,family = 'binomial')
glm2 <- glm(cursmoke ~ age+sex+diabetes, data=smoke_vs1,family = 'binomial')
a=(summary(glm1)$coefficients[2])-(1.96 * (summary(glm1)$coefficients[5])) 
b= (summary(glm1)$coefficients[2])+(1.96 * (summary(glm1)$coefficients[5]))
c=summary(glm2)$coefficients[2]
exp(a)
exp(b)
exp(c)
!(exp(c)>=exp(a) & exp(c)<=exp(b))
d=(summary(glm1)$coefficients[3])-(1.96 * (summary(glm1)$coefficients[6]))
e=(summary(glm1)$coefficients[3])+(1.96 * (summary(glm1)$coefficients[6]))
f=summary(glm2)$coefficients[3]
exp(d)
exp(e)
exp(f)
!(exp(f)>=exp(d) & exp(f)<=exp(e))
```
diabetes is not confounder for sex and age with cursmoke

None of these are confounders. It makes sense as nothing could affect age. Same for question 2. So we only put age and sex into the model.

```{r}
smoke_vs3 = smoke %>% 
  dplyr::select(c(randid,sex,age,cursmoke)) %>% 
  mutate(cursmoke = as.factor(cursmoke), sex=as.factor(sex)) %>% 
  na.omit()
glmer_1 <- glmer(cursmoke ~ age *sex + (1 | randid),
                 data = smoke_vs3, 
                 family = binomial)
summary(glmer_1) 
```
significant. 


```{r}
#CI age
-0.23847-1.96*0.01456
-0.23847+1.96*0.01456
```


```{r}
#CI sex
-6.93043-1.96*1.09252
-6.93043+1.96*1.09252
```

```{r}
#CIage:sex2     
0.05611-1.96*0.01671
0.05611+1.96*0.01671
```


#Question 2
When looking at cigarette packs smoked per day, it appears that the number steadily decreases as individuals get older. The trend once again is the same in each sex however females are smoking less packs a day overall.

```{r}

smoke %>%
  ggplot(aes(age, cigpday)) +
  geom_smooth(method = "loess") +
  ggtitle("Figure 3: Cigarette per Day across Age")

smoke %>%
  mutate(sex = as.factor(sex)) %>%
  ggplot(aes(age, cigpday, group = sex, color = sex)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Figure 4: Cigarettes per Day across Age") 
  
```



Variable Selection and confounder identidication:

```{r}
smoke_vs2 = smoke %>% 
  filter(period==1) %>% 
  dplyr::select(c(randid,sex,age,cigpday,totchol,bmi,heartrte,educ,diabp,sysbp,diabetes)) %>% 
  mutate(sex=as.factor(sex), diabetes=as.factor(diabetes)) %>% 
  na.omit()
```

```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+totchol, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
totchol is not confounder for age and sex with cigpday

```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+bmi, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
bmi is not confounder for age and sex with cigpday


```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+heartrte, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
heartrte is not confounder for age and sex with cigpday


```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+educ, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
educ is not confounder for age and sex with cigpday

```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+diabp, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
diabp is not confounder for age and sex with cigpday
```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+sysbp, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
sysbp is not confounder for age and sex with cigpday


```{r}
lm1 <- lm(cigpday ~ age+sex, data=smoke_vs2)
lm2 <- lm(cigpday ~ age+sex+diabetes, data=smoke_vs2)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[5])) 
b= (summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[5]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c=a & c<=b)
d=(summary(lm1)$coefficients[3])-(1.96 * (summary(lm1)$coefficients[6]))
e=(summary(lm1)$coefficients[3])+(1.96 * (summary(lm1)$coefficients[6]))
f=summary(lm2)$coefficients[3]
d
e
f
!(f>=d & f<=e)
```
diabetes is not confounder for age and sex with cigpday


```{r}
smoke_vs4 = smoke %>% 
  dplyr::select(c(randid,cigpday,sex,age)) %>% 
  mutate(sex=as.factor(sex)) %>% 
  na.omit()

smoke_vs4_nonsmoker = smoke_vs4 %>% filter(cigpday == 0) %>% group_by(randid) %>%
summarize(cig_count = sum(cigpday)) %>% filter(cig_count == 0)
nonsmoker_id = unique(smoke_vs4_nonsmoker$randid)
smoke_vs4_smoker = smoke_vs4 %>% filter(!randid %in% nonsmoker_id)
```

```{r}
lmer_2 <- lmer(cigpday ~ age * sex + (1 | randid),
                    data = smoke_vs4_smoker)
summary(lmer_2)
 
```
Ignore those who did not smoke through the whole study.

pvalue
```{r}
coefs2 <- data.frame(coef(summary(lmer_2)))
# use normal distribution to approximate p-value
coefs2$p_value <- 2 * (1 - pnorm(abs(coefs2$t.value)))
coefs2
```

```{r}
#CI age
-0.06629-1.96*0.02622
-0.06629+1.96*0.02622
```

```{r}
#CI sex
-15.59130-1.96*2.00555
-15.59130+1.96*2.00555
```

```{r}
#CI age:sex2      
0.17357-1.96*0.03739
0.17357+1.96*0.03739
```

# Part 2
In Figure 5, we see that as systolic blood pressure increases the likelihood of smoking decreases. The trend is not as profound in Figure 6 with diastolic BP or with serum total cholesterol in Figure 7. 

```{r}

smoke %>%
  mutate(cursmoke = as.factor(cursmoke)) %>%
  ggplot(aes(age ,sysbp, group = cursmoke, color = cursmoke)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Figure 5: Systolic Blood Pressure across Age")

smoke %>%
  mutate(cursmoke = as.factor(cursmoke)) %>%
  ggplot(aes(age, diabp, group = cursmoke, color = cursmoke)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Figure 6: Diastolic Blood Pressure across Age")

smoke %>%
  mutate(cursmoke = as.factor(cursmoke)) %>%
  ggplot(aes(age, totchol, cursmoke, group = cursmoke, color = cursmoke)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Figure 7: Total Cholesterol across Age")
```

#Question 3

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + sex, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
sex is not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + age, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
age is confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + totchol, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
totchol is not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + bmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
bmi is confounder



```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + heartrte, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
heartrte is not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + educ, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
educ is not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + diabetes, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
diabetes is not confounder


```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + prevap, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
prevap is not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + prevchd, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + prevmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(sysbp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(sysbp ~ cursmoke + prevstrk, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

By rule of thumb, age and bmi are confounders. Based on the literature view, sex can be potential confounders as they can affect smoke status and sysbp at the same time. So we still put it into model.

```{r}
smoke_vs5 = smoke %>% 
  dplyr::select(c(randid,cursmoke,sex,age,bmi,sysbp)) %>% 
  mutate(sex=as.factor(sex),cursmoke=as.factor(cursmoke)) %>% 
  na.omit()
```


```{r}

lmer_3 <- lmer(sysbp ~ cursmoke + bmi + sex + age + (1|randid), data = smoke_vs5)
summary(lmer_3)
```
Calculating p value using normal approximation:

```{r}
coefs3 <- data.frame(coef(summary(lmer_3)))
# use normal distribution to approximate p-value
coefs3$p_value <- 2 * (1 - pnorm(abs(coefs3$t.value)))
coefs3
```

No interaction term because not significant

Cursmoke not significant,but include because this is our interest
```{r}
#cursmoke 95% CI 
-0.008418-1.96*0.432847
-0.008418+1.96*0.432847
```

```{r}
#CI bmi
1.443149-1.96*0.057512
1.443149+1.96*0.057512
```


```{r}
#95%CI sex
2.609485-1.96*0.552576
2.609485+1.96*0.552576
```


```{r}
#age
0.855104-1.96*0.020583
0.855104+1.96*0.020583
```



#Question 4
```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + sex, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
sex is not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + age, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
age is confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + totchol, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
totchol is not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + bmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
bmi is confounder



```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + heartrte, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
heartrte is not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + educ, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
educ is not confounder


```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + diabetes, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
diabetes is not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + prevap, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + prevchd, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + prevmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(diabp ~ cursmoke, data=smoke_vs1)
lm2 <- lm(diabp ~ cursmoke + prevstrk, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

By rule of thumb, age, bmi is confounder. Based on the literature view, sex can be potential confounders as it can affect smoke status and diabp at the same time. So we still put them into model.

checking interactions
```{r}
smoke_vs6 = smoke %>% 
  dplyr::select(c(randid,cursmoke,sex,age,bmi,diabp)) %>% 
  mutate(sex=as.factor(sex),cursmoke=as.factor(cursmoke)) %>% 
  na.omit()
smoke_vs6 %>%
  mutate(cursmoke = as.factor(cursmoke)) %>%
  ggplot(aes(bmi ,diabp, group = cursmoke, color = cursmoke)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("Systolic Blood Pressure across bmi")
```

This plot argues for some interaction of the two predictors, as the lines are not parallel.

```{r}

lmer_4 <- lmer(diabp ~  cursmoke*bmi + sex  +cursmoke*age + (1|randid), data = smoke_vs6)
summary(lmer_4)
```
pvalue

```{r}
coefs4 <- data.frame(coef(summary(lmer_4)))
# use normal distribution to approximate p-value
coefs4$p_value <- 2 * (1 - pnorm(abs(coefs4$t.value)))
coefs4
```
cursmoke1:bmi not siginificant, remove this iteraction

```{r}

lmer_41 <- lmer(diabp ~  bmi + sex  +cursmoke*age + (1|randid), data = smoke_vs6)
summary(lmer_41)
```

```{r}
coefs41 <- data.frame(coef(summary(lmer_41)))
# use normal distribution to approximate p-value
coefs41$p_value <- 2 * (1 - pnorm(abs(coefs41$t.value)))
coefs41
```


```{r}
#CI cursmoke
-8.81416-1.96*1.17670
-8.81416+1.96*1.17670
```

```{r}
#CIbmi
1.02005-1.96*0.03139
1.02005+1.96*0.03139
```

sex although not significant, it is confounder, so we still put in into the model
```{r}
#CI sex
-0.53287-1.96*0.29288
-0.53287+1.96*0.29288
```

```{r}
#CI age
-0.05110-1.96*0.01442
-0.05110+1.96*0.01442
```

```{r}
#cursmoke1:age  
0.16118-1.96*0.02129
0.16118+1.96*0.02129
```

#QUestion 5
```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + sex, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
sex is not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + age, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
age is confounder



```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + diabp, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
diabp is not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + bmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
bmi is confounder


```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + heartrte, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
heartrte is not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + educ, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
educ is not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + sysbp, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
sysbp is not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + diabetes, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
diabetes is not confounder


```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + prevap, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + prevchd, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + prevmi, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + prevstrk, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

```{r}
lm1 <- lm(totchol ~ cursmoke, data=smoke_vs1)
lm2 <- lm(totchol ~ cursmoke + prevhyp, data=smoke_vs1)
a=(summary(lm1)$coefficients[2])-(1.96 * (summary(lm1)$coefficients[4])) 
b=(summary(lm1)$coefficients[2])+(1.96 * (summary(lm1)$coefficients[4]))
c=summary(lm2)$coefficients[2]
a
b
c
!(c>=a & c<=b)
```
not confounder

By rule of thumb, age, bmi is confounder. Based on the literature view, sex can be potential confounders as it can affect smoke status and totchol at the same time. So we still put them into model.



```{r}
smoke_vs7 = smoke %>% 
  dplyr::select(c(randid,cursmoke,sex,age,bmi,totchol)) %>% 
  mutate(sex=as.factor(sex),cursmoke=as.factor(cursmoke)) %>% 
  na.omit()
```

checking interactions
```{r}
smoke_vs7 %>%
  mutate(cursmoke = as.factor(cursmoke)) %>%
  ggplot(aes(bmi ,totchol, group = cursmoke, color = cursmoke)) +
  geom_smooth(method = "loess", se = F) +
  ggtitle("totchol across bmi")
```
This plot argues for some interaction of the two predictors, as the lines are not parallel.


```{r}

lmer_5 <- lmer(totchol ~ cursmoke*bmi + sex +cursmoke*age + (1|randid), data = smoke_vs7)
summary(lmer_5)
```

```{r}
coefs5 <- data.frame(coef(summary(lmer_5)))
# use normal distribution to approximate p-value
coefs5$p_value <- 2 * (1 - pnorm(abs(coefs5$t.value)))
coefs5
```


```{r}
#CI cursmoke
-29.38692-1.96*7.05633
-29.38692+1.96*7.05633
```

```{r}
#CI bmi
1.45643-1.96*0.15257
1.45643+1.96*0.15257
```

```{r}
#CI sex
13.94202-1.96*1.22545
13.94202+1.96*1.22545
```



```{r}
#CI age
0.32861-1.96*0.05514
0.32861+1.96*0.05514
```


```{r}
#CI cursmoke1:bmi   
0.74739-1.96*0.21240
0.74739+1.96*0.21240
```


```{r}
#CI  cursmoke1:age   
0.26381-1.96*0.08053
0.26381+1.96*0.08053
```
Please include a table which shows point estimate, 95 CI and p value for each term in the model(calculated above)